Preregistration for modelling in ecolgy and conservation

### 2020-05-26 index.Rmd
#20 update manuscript - write UX design section

---
title: "EcoConsPreReg"
description: |
  A new article created using the Distill format.
author:
  - name: Elise Gould 
    affiliation: School of BioSciences, University of Melbourne
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    self_contained: false
---
The Abstract must not exceed 350 words and should list the main results and conclusions, using simple, factual, numbered statements:

@@

**Keywords: preregistration, preregistration template, modelling, applied ecology, conservation decision-making, metaresearch, reproducibility, non-hypothesis testing(?)**

@@

\*\*About the Approach\*\*

As Pu et al. (2019) identify, a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings. In order to avert this mismatch, we have taken a user-centred design approach to the problem of preregistration. We do this in two ways, firstly, by running a series of structured and collaborative activities in a workshop setting to reflect on our own research experiences in order to capture scientific workflows as they occur practice. And secondly to, use a real-world research problem as a case-study to test and evaluate the templates for utility and feasibility in practice.

\*\*Defining the purpose of preregistration\*\*

Pu et al's key finding:

Pu et al. "Current designs represent a failure of designers to make their internal model of preregistration clear to users: the purpose of each question, and indeed the entire form, is often opaque."

Pu et al found that because different users ascribed different purposes to preregistration, it resulted in conflicting designs of preregistration formats, which they argue may ultimately undermine the overarching goal of preregistration: "improving the reliability of scientific findings".

We state the intended purposes of the preregistration template up front:

\*\*Acknowledging different use-cases\*\*

Pu et al:

\- norms around how people use preregistration are uncertain, and vary among users.

Our use-cases:

1\. As creators

2\. As users:

@@

@@

>>>>

---
title: "Preregistration for modelling in ecolgy and conservation"
description: |
  Working Draft.
author:
  - name: Elise Gould 
    affiliation: School of BioSciences, University of Melbourne
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    self_contained: false
---

Intended journal: Methods in Ecology and Evolution.

The Abstract must not exceed 350 words and should list the main results and conclusions, using simple, factual, numbered statements:

@@

**Keywords: preregistration, preregistration template, modelling, applied ecology, conservation decision-making, metaresearch, reproducibility, non-hypothesis testing(?) (no more than 8 for MEE).**

@@

**Designing Preregistration Templates with a User-Centred Design Approach**

As Pu et al. (2019) identify, a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings. In order to avert this mismatch, we have taken a user-centred design approach to the problem of preregistration. We do this in two ways, firstly, by running a series of structured and collaborative activities in a workshop setting to reflect on our own research experiences in order to capture scientific workflows as they occur practice. And secondly, by using a real-world research problem as a case-study to test and evaluate the templates for utility and feasibility in practice.

@TODO what does a UX design approach entail... stating purpose, use-case and contexts a priori? Explain.

**Intended Purpose of this Preregistration**

Defining the purpose and intended use-cases of a preregistration template has important implications for the design of the template --- from its structure, content, user-interface and use within the broader research workflow. We define the primary purpose of the template as "to delimit researcher degrees of freedom" when conducting model-based research in applied ecology and conservation. Although others have defined the purpose of preregistration to include 'delineating between *exploratory* from *confirmatory* research' \<refs\>, this is not the intended purpose of the template here. Beyond hypothesis-testing contexts, this is a simplifying dichotomy that, in practice, is more blurred and nuanced than the ideal. For instance researchers might move back and forth between different research-modes throughout the modelling cycle. Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, perhaps with the aid of some philosophers of science. Consequently, we wish to deemphasise the purpose of distinguishing between exploratory and confirmatory research, and focus the design of the templates in being able to delimit researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and other questionable research practices, like cherry-picking, HARKing etc. \<insert ref to my QRP manuscript\>.

The secondary purpose of the model is to increase *research transparency* of model-based research in ecology and conservation. *Research transparency* refers to both production transparency --- which includes research artefacts like open-access data and materials, or data collection procedures --- and analytic transparency --- which describes a complete account of how the analytic conclusions are drawn from the data (Lupia and Elman, 2014). \<refs to model reporting literature- it's dire, can't properly be evaluated by a reviewer, moreover can't be reproduced\>

**Preregistration template users and research-contexts**

Given the stated purposes of the preregistration templates we designed, it implies two types of users: creators of preregistrations using the template, and reviewers of a completed template, including reviewers, journal editors and any readership. The content and structure of the items on the template will be designed with the preregistration-creator and the preregistration readership in mind - it will capture critical decision-points that may influence the overall model and analysis and recommendations or decisions that the model informs, thus delimiting researcher degrees of freedom at these junctures in the modelling decision process. It should also enhance the research transparency of the study, ensuring that these choices are documented with sufficient transparency to let the reader make informed decisions about the credibility and reliability of the model and associated analysis and interventions.

More specifically, we intend this template to be suitable not just in 'pure-research'-only contexts, but to be suitable for ecologists and conservation decision-makers who are working in the 'actionable sciences' \<which guy was this again.. Sutherland?? define\>.  The template should be adequate for use by scientists and modellers working in applied ecological contexts, but also for management and conservation decision-making contest where there are real-world decisions hinging on the model-based research.

\<Research Types: Modelling - what do we mean by modelling?? I think if we don't specify this then people will just think 'classical linear regression' as modelling, and then will say, well why don't use existing hypothesis testing templates, like AsPredicted. I need to specify some particularities about the type of modelling that occurs in in these research contexts, and why existing templates don't suit, and that we need to design specifically for these research contexts.\>

**Use-Cases**

Depending on what role is assumed, there are different 'use-cases' for authors, reviewers/editors *Pu et al.* However, these use-cases are not typically explicitly designed for.  For example, reviewers and editors (and sometimes authors) face the problem of 'checking' --- that is, 'the process of comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified.' Existing preregistration formats are designed with the role of the author in mind --- they are text-based documents, that are ill-suited to checking the reported study against the preregistration.

Consequently, we aim not just to simply design a new template, deciding on its structure and content, but we propose a workflow and methodology to be implemented by the author when completing a preregistration of their model-based research \<see section - inner logic of the model - discuss features model-based research, like iterative development?\>

@@

What attributes do we aim for the template

A) an interface that ensures ease of use for the creator of a preregistration

B) structure and content that is necessarily opinionated.. forces user to a particular workflow, so getting that workflow right, that reflects both norms and practice of model-based research is pivotal.

@@

@TODO Insert screenshot of app.

@TODO link to shiny app.

As *Pu et al* find, norms around preregistration are both uncertain and vary among users. By norms, we mean \<insert definition from Pu\>. Although the design of the template in terms of the content and structure is *opinionated*, in that it forces a particular view of the modelling research workflow, people are free to modify the template according to their needs. \<The preregistration infrastructure we provide here allows for people to modify the template easily - set up the infrastructure of the shiny app to take in a CSV file, which is then exported to a JSON file for creating the shiny app (alt, just modify the JSON directly)\>
